---
title: "SABR: Top ASVs correlated with g N (N2O) per hectare"
author:
  - "Bolívar Aponte Rolón"
date: last-modified
format:
#  html:
#    theme: cosmo
#    toc: true
#    toc-depth: 3
#    number-sections: true
#    code-fold: true
#    df-print: paged
  pdf:
    documentclass: scrartcl
    papersize: letter    
    fig-width: 7.5            # fits across page
    fig-asp: 0.8
    fig-align: center

execute:
  echo: true
  warning: false
  message: false
  cache: false
  freeze: auto
---

# Data preparation and goal

**Goal**: end up with 20 ASVs per prevalence group (e.g., prevalent_80), across both sampling locations combined, not 20 per sampling location (e.g. 20 at "interrow" and 20 at "row").

### *Pre-filtering: Find the best ASVs within each location*

For each prevalence group and each sampling location separately (e.g., Row and interrow), we: 1) Compute the *mean relative abundance* of every ASV.

2)  Within each prevalence × location, we take the top 20 ASVs by that mean.

**Result**: a shortlist that may include up to 40 ASVs per prevalence (20 from Row + 20 from interrow), with duplicates if the same ASV is top in both locations.

### *Pre-filtering: For each ASV within a prevalence group*

Combine locations and re-rank to keep only 20 per prevalence

1)  Collapse that shortlist to unique ASVs per prevalence (i.e., remove duplicates across locations).

2)  Average its “top-list” means across locations to get one overall score.

3)  Re-rank ASVs within each prevalence group by this overall score and keep the top 20.

This ensured the top 20 ASVs per prevalence group, even if *row* and *interrow* had different top-20s.

## Summary

- **Consistency:** The top ASVs display recurring trends across prevalence thresholds.

- **Significance:** Few row–interrow contrasts are statistically significant, despite visible patterns.

- **Directionality:** Several ASVs switch direction by location (positive in row, negative in interrow).

- **Magnitude:** Many ASVs show steeper slopes, indicating stronger responses to $[g, N(N2O)ha−1]$, even when not statistically significant.

The ASVs that show inverse relationships across locations maybe consistent with microenvironment- or management-specific responses. 

::: callout-note
Render from the repository root:

-   CLI: `quarto render sabr_top_asvs_nitrogen.qmd --execute-dir .`
-   R: `quarto::quarto_render("sabr_top_asvs_nitrogen.qmd")`
:::

```{r}
#| label: setup
#| include: false
# Keep objects in the global knit environment so both scripts can share state if needed
knitr::knit_hooks$set(document = function(x) x)
# sets the working directory for code execution to the repo root
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
source(
  "R/005_core_microbiome.R",
  local = knitr::knit_global(),
  echo = TRUE,
  print.eval = TRUE
)
```