---
title: "SABR: Top ASVs correlated with g N (N2O) per hectare"
author:
  - "Bolívar Aponte Rolón"
date: last-modified
format:
#  html:
#    theme: cosmo
#    toc: true
#    toc-depth: 3
#    number-sections: true
#    code-fold: true
#    df-print: paged
  pdf:
    documentclass: scrartcl
    papersize: letter

execute:
  echo: true
  warning: false
  message: false
  cache: false
  freeze: auto
---

# Data preparation and goal

**Goal**: end up with 20 ASVs per prevalence group (e.g., prevalent_80), across both sampling locations combined, not 20 per sampling location (e.g. 20 at "interrow" and 20 at "row").

_Pre-filtering: Find the best ASVs within each location_
For each prevalence group and each sampling location separately (e.g., Row and interrow), we:
1) Compute the _mean relative abundance_ of every ASV.

2) Within each prevalence × location, we take the top 20 ASVs by that mean.

**Result**: a shortlist that may include up to 40 ASVs per prevalence (20 from Row + 20 from interrow), with duplicates if the same ASV is top in both locations.

_For each ASV (within a prevalence group)_
Combine locations and re-rank to keep only 20 per prevalence

1) Collapse that shortlist to unique ASVs per prevalence (i.e., remove duplicates across locations).

2) Average its “top-list” means across locations to get one overall score.

3) Re-rank ASVs within each prevalence group by this overall score and keep the top 20.

This ensured the top 20 ASVs per prevalence group, even if Row and interrow had different top-20s.


::: callout-note
Render from the repository root:

-   CLI: `quarto render sabr_top_asvs_nitrogen.qmd --execute-dir .`
-   R: `quarto::quarto_render("sabr_top_asvs_nitrogen.qmd")`
:::

```{r}
#| label: setup
#| include: false
# Keep objects in the global knit environment so both scripts can share state if needed
knitr::knit_hooks$set(document = function(x) x)
# sets the working directory for code execution to the repo root
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
source(
  "R/005_core_microbiome.R",
  local = knitr::knit_global(),
  echo = TRUE,
  print.eval = TRUE
)
```